---
title: "Speaker WebPPL Model"
author: "Ben Morris and Dan Yurovsky"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    theme: flatly
    toc: true
    toc_float:
      collapsed: true
    code_folding: hide
---


### Setup
```{r, include=FALSE}
knitr::opts_chunk$set(fig.align='center', echo=FALSE, messages=FALSE, warning = FALSE,
                      fig.height = 3, fig.width=5)

#Load packages + setup
library(rwebppl)
library(tidyverse)
library(ggjoy)
library(directlabels)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

theme_set(theme_bw(base_size = 14) + theme(panel.grid = element_blank(),
                  strip.background = element_blank()))
```


### Run Model (or read in saved copy below)
```{r}
library(RCurl)
full_implicit <- read.csv(text =getURL("https://rawgit.com/benjamincmorris/reference-game/master/turk/2017.5.9_sender_full/5.9_data_anon.csv"))

all_data_dropped <- full_implicit %>% filter(toBeDropped==0)

tmp <- all_data_dropped %>% group_by(speed, method) %>%
  filter(duration<15000) %>%
  filter(responseCorrect==1) %>%
  mutate(hyp_score= ceiling(11-(duration*2/1000))*10) %>%
  summarize(medDur = median(duration), sdDur = sd(duration),
            meanScore=mean(score), n =n(), avgCorr = mean(responseCorrect),
            meanHypScore = mean(hyp_score)) 

implicit_points <- tmp %>%
  select(speed, method, meanScore, meanHypScore) %>%
  gather(type, Score, meanScore, meanHypScore) %>%
  spread(method, Score)

tmp %>%
  gather(type, Score, meanScore, meanHypScore) %>%
  mutate(Score = if_else(Score < 0, 0, Score))  %>%
  mutate(method=factor(method, levels=c("label","click", "label_click"))) %>%
  ggplot(aes(x=method))+
  geom_bar(stat="identity", aes(x=method, y=Score, fill=method)) +
  # geom_bar(aes(x=method, y=medDur))+
  facet_grid(type~speed) +
  theme_bw()
#100_30 would translate to about 65 speed

empiricalImplicitVocabs <- full_implicit %>%
  group_by(ldf_num, speed, partnersExposure, realLabel, exposureRate) %>%
  summarize(testCorrect=first(testCorrect)) %>%
  left_join(implicit_points) %>%
  rename(condition = speed,
         exposures=exposureRate,
         label_points=label,
         click_points= click,
         label= realLabel) %>%
  ungroup() %>%
  mutate(known = ifelse(testCorrect == 1, TRUE, FALSE),
         partnersExposure = ifelse(partnersExposure=='Perfect', 2, 
                                   ifelse(partnersExposure=='None', 0, .5)),
         condition=as.factor(condition))  %>%
  as.data.frame()
  
 
 
 
# ##read in empirical data to feed model
# empiricalVocabsConds<- read.csv("empiricalVocabsFull.csv") %>%
  # select(-X)
# # ### this line runs webppl model
t1 <- proc.time()
# outcomes<-webppl(program_file = "speaker.wppl", data=empiricalVocabsConds, data_var = "empiricalVocabs")
# outcomes_implicit<-webppl(program_file = "speaker.wppl", data=empiricalImplicitVocabs, data_var = "empiricalVocabs")
outcomes_implicit_luce4<-webppl(program_file = "speaker.wppl", data=empiricalImplicitVocabs, data_var = "empiricalVocabs")

proc.time() - t1
# # 
# # ### clean up data from model
gamePredictions_luce4 <- outcomes_implicit_luce4 %>%
  unnest(predictions, pred_perf_teaching, gameTrials) %>%
  unnest(pred_perf_teaching, gameTrials) %>%
    cbind(appearance= rep.int(c(1,2,3), nrow(.)/3))



```


### Read In and Join Empirical Data
```{r}
###full empirical dataset
# full_empirical <- read.csv('empirical_data/5.12_data_anon.csv') %>% select(-X)
# 
# ###join empirical data and model predictions
all_data_luce4 <- full_implicit %>%
  filter(ldf_num %in% gamePredictions_luce4$me) %>%
  select(-partnersExposure) %>%
  filter(toBeDropped != 1) %>%
  group_by(ldf_num, exposureRate, targetObjectName) %>%
  mutate(appearance = if_else(trialnum == min(trialnum), 1,
                              if_else(trialnum == max(trialnum), 3, 2))) %>%
  left_join(gamePredictions_luce4, by=c('ldf_num'='me', 'realLabel'='gameTrials', 'appearance')) %>%
  mutate(method=ifelse(method=='click', "point",
                        ifelse(method=="label","speak", "teach"))) %>%
  as.data.frame(.)

```


#Or read in all data (inc. model predictions) from saved copy
```{r}
###read in model data from csv
all_data <- read.csv("5.12_three_models.csv") %>% select(-X)
# all_data <- read.csv("fixed_teach_cost.csv") %>% select(-X)

```


#Plot 1: Method Choice as a function of Appearance, split by Utility Condition
```{r}
seprop <- function(props) {
  mean_prop = mean(props)
  sqrt( (mean_prop*(1 - mean_prop)) / length(props))
}


modality.colors <- c(speak = "#e8250b", point = "#1f11e0", teach ="#54a832", speech = "#e8250b", gesture = "#1f11e0", both ="#54a832")



###wrangling data for plots by exposure
prop_methods_exposures <- all_data %>%
  filter(toBeDropped != 1) %>%
  # filter(partnersExposure == 0) %>%
  ungroup() %>%
  select(condition,ldf_num, partnersExposure, exposureRate,method, pred_perf_teaching) %>%
  gather(isEmpirical, method, -condition, -partnersExposure, -ldf_num, -exposureRate, pred_perf_teaching) %>%
  group_by(condition, ldf_num,partnersExposure, exposureRate,isEmpirical, method) %>%
  summarise(n = n()) %>%
  group_by(condition, ldf_num,partnersExposure, isEmpirical, exposureRate) %>%
  mutate(n = n/sum(n)) %>%
  ungroup() %>%
  mutate(method = as.factor(method)) %>%
  tidyr::complete(nesting(condition, ldf_num, partnersExposure), isEmpirical, exposureRate, method, fill = list(n = 0)) %>%
  group_by(condition, partnersExposure, method, isEmpirical, exposureRate) %>%
  summarise(mean = mean(n), se = seprop(n)) %>%
  ungroup() %>%
  mutate(condition=factor(condition,labels=c('Talk is Cheap', 'Talk is Less Cheap'))) %>%
  mutate(exposureRate=as.factor(exposureRate)) %>%
  filter(method!='teach') %>%
  filter(partnersExposure==.5)

prop_methods_exposures %>% 
  filter(isEmpirical=="method") %>%
  ggplot(aes(x=exposureRate, y=mean)) +
  geom_point(aes(color=method),size=1.5,position=position_dodge(.25)) +
  ## plotting model predictions
  # geom_ribbon(data=(prop_methods_exposures %>% filter(isEmpirical=="pred_perf_teaching")),
  #   aes(x=as.numeric(exposureRate),ymax = mean + se, ymin = mean - se,
  #       fill=method, alpha=.4), position=position_dodge(.25)) +
  geom_line(aes(x=(exposureRate), y=mean, group=interaction(method,condition), color=method),  size=.9,
          position=position_dodge(.25)) +
  geom_linerange(aes(ymax = mean + se,
    ymin = mean - se, color=method), position=position_dodge(.25)) +
  # facet_grid(partnersExposure ~ condition) +
  labs(y="Proportion of Trials", x="Exposure Rate during Training") +
  coord_cartesian(ylim=c(0,1)) +
  facet_wrap(~condition) +
  # scale_alpha(range = c(0.4, 0.8)) +
  theme_bw() +
  scale_color_manual(values=modality.colors,
    name = "Modality", labels=c("Gesture", "Speech")) +
  scale_fill_manual(values=modality.colors,
    name = "Modality", labels=c("Gesture", "Speech"))


###plot of method choice as a function of appearance, split across point conditions
# to_be_plotted <- prop_methods %>% 
#   filter(partnersExposure==.5) %>%
#   filter(method!='teach')
# 
# to_be_plotted %>% filter(isEmpirical=='method') %>%
#   ggplot(aes(x=appearance, y=mean)) +
#   geom_line(aes(x=as.factor(appearance), y=mean, group=interaction(method,condition), color=method, shape=condition), position=position_dodge(.25)) +
#   geom_pointrange(aes(ymax = mean + se,
#     ymin = mean - se, color=method, alpha=condition), position=position_dodge(.25)) +
#   # facet_grid(partnersExposure ~ condition) +
#   labs(y="Proportion of Trials", x="Object Appearance #") +
#   coord_cartesian(ylim=c(0,1)) +
#   # geom_ribbon(data=(to_be_plotted %>% filter(isEmpirical=="pred_perf_teaching")),
#               # aes(x=as.numeric(appearance),ymax = mean + se,
#                       # ymin = mean - se, fill=method, alpha=.4), position=position_dodge(.25)) +
#   # facet_wrap(~condition) +
#     scale_fill_brewer(name = "Condition", 
#                     labels=c("Talk is Cheap", "Talk is More Expensive"),
#                     palette = "Set1") +
#   theme_bw()


```


#Teaching Plot Across Conditions
```{r}
###wrangling data for teaching
prop_methods_teaching <- all_data %>%
  filter(toBeDropped != 1) %>%
  # filter(partnersExposure == 0) %>%
  ungroup() %>%
  select(condition,ldf_num, partnersExposure, appearance,method,predictions, pred_perf_partner, pred_perf_teaching) %>%
  # select(condition,ldf_num, partnersExposure, appearance,method) %>%
  gather(isEmpirical, method, -condition, -partnersExposure, -ldf_num, -appearance) %>%
  group_by(condition, ldf_num,partnersExposure, appearance,isEmpirical, method) %>%
  summarise(n = n()) %>%
  group_by(condition, ldf_num,partnersExposure, isEmpirical, appearance) %>%
  mutate(n = n/sum(n)) %>%
  ungroup() %>%
  mutate(method = as.factor(method)) %>%
  tidyr::complete(nesting(condition, ldf_num, partnersExposure), isEmpirical, appearance, method, fill = list(n = 0)) %>%
  group_by(condition, partnersExposure, method, isEmpirical, appearance) %>%
  summarise(mean = mean(n), se = seprop(n)) %>%
  filter(method=='teach', condition=="100_30") %>%
  ungroup() %>%
  mutate(partnersExposure=factor(partnersExposure, labels=c('None', 'Half', 'Twice'))) %>%
  mutate(appearance=factor(appearance, labels=c('First','Second', 'Third')))

prop_methods_teaching %>%
  filter( isEmpirical=='method') %>%
  ggplot(aes(x=appearance, y=mean, label=partnersExposure)) +
  geom_point(aes(x=appearance, y=mean, color=partnersExposure), size=2.25, position=position_dodge(.25)) +
  # geom_ribbon(data=(prop_methods_teaching %>% filter(isEmpirical=="pred_perf_teaching")),
  #   aes(x=as.numeric(appearance),ymax = mean + se, ymin = mean - se, 
  #       fill=partnersExposure, alpha=.4), position=position_dodge(.25)) +
  geom_line(aes(x=appearance, y=mean, group=partnersExposure, color=partnersExposure), 
            size=1, position=position_dodge(.25)) +
  geom_linerange(aes(ymax = mean + se,
    ymin = mean - se, color=partnersExposure), position=position_dodge(.25)) +
  # facet_grid(partnersExposure ~ condition) +
  labs(y="Proportion of Teaching Trials", x="Object Instance during Game") +
  coord_cartesian(ylim=c(0,.6)) +
  geom_dl(aes(label = partnersExposure, color=partnersExposure), method=list("first.qp", hjust=c(.95, 1.45, 1.55), cex = .9)) +
  # scale_color_manual(values=c("#87d868", "#54a832", "#2c6d12"), name=NULL, labels=NULL) 
  scale_color_manual(values=c("#87d868", "#54a832", "#2c6d12"), name = "Partner's Exposure", 
                    labels=c("None", "Half as Much", "Twice as Much")) +
  scale_fill_manual(values=c("#87d868", "#54a832", "#2c6d12"), name = "Partner's Exposure", 
                    labels=c("None", "Half as Much", "Twice as Much")) 
```



#Teaching Plot by Exposure Rate
###First Trial ONLY
```{r}
###wrangling data for teaching
prop_methods_exposures <- all_data %>%
  filter(toBeDropped != 1) %>%
  filter(appearance == 1) %>%
  ungroup() %>%
  # select(condition,ldf_num, partnersExposure, exposureRate,method,predictions, pred_perf_partner, pred_perf_teaching) %>%
  select(condition,ldf_num, partnersExposure, exposureRate,method) %>%
  gather(isEmpirical, method, -condition, -partnersExposure, -ldf_num, -exposureRate) %>%
  group_by(condition, ldf_num,partnersExposure, exposureRate,isEmpirical, method) %>%
  summarise(n = n()) %>%
  group_by(condition, ldf_num,partnersExposure, isEmpirical, exposureRate) %>%
  mutate(n = n/sum(n)) %>%
  ungroup() %>%
  mutate(method = as.factor(method)) %>%
  tidyr::complete(nesting(condition, ldf_num, partnersExposure), isEmpirical, exposureRate, method, fill = list(n = 0)) %>%
  group_by(condition, partnersExposure, method, isEmpirical, exposureRate) %>%
  summarise(mean = mean(n), se = seprop(n))

prop_methods_exposures %>%
  filter(method=='teach', isEmpirical=='method', condition=="100_30") %>%
  ungroup() %>%
  mutate(partnersExposure=factor(partnersExposure, labels=c('None', 'Half', 'Twice'))) %>%
  mutate(exposureRate=as.factor(exposureRate)) %>%
  ggplot(aes(x=exposureRate, y=mean, label=partnersExposure)) +
  geom_point(aes(x=exposureRate, y=mean, color=partnersExposure), size=2.25, position=position_dodge(.25)) +
  geom_line(aes(x=exposureRate, y=mean, group=partnersExposure, color=partnersExposure), size=1, position=position_dodge(.25)) +
  geom_linerange(aes(ymax = mean + se,
    ymin = mean - se, color=partnersExposure), position=position_dodge(.25)) +
  # facet_grid(partnersExposure ~ condition) +
  labs(y="Teaching Trials (Proportion)", x="Exposure Rate") +
  coord_cartesian(ylim=c(0,.75)) +
  geom_dl(aes(label = partnersExposure, color=partnersExposure), method=list("first.qp", hjust=c(1.05, 1.45, 1.5), cex = .9)) +
  scale_color_manual(values=c("#87d868", "#54a832", "#2c6d12"), name = "Partner's Exposure", 
                    labels=c("None", "Half as Much", "Twice as Much")) 
  # geom_ribbon(data=(to_be_plotted %>% filter(isEmpirical=="pred_perf_teaching")),
              # aes(x=as.numeric(appearance),ymax = mean + se,
                      # ymin = mean - se, fill=method, alpha=.4), position=position_dodge(.25)) +
  # facet_wrap(~condition) 
```





#Plot of Model Predictions vs. Empirical Data
```{r}
prop_methods_luce4 <- all_data_luce4 %>%
  filter(toBeDropped != 1) %>%
  rename(condition=speed) %>%
  # filter(partnersExposure == 0) %>%
  ungroup() %>%
  select(condition,ldf_num, partnersExposure, appearance,method,predictions,  pred_perf_teaching) %>%
  # select(condition,ldf_num, partnersExposure, appearance,method) %>%
  gather(isEmpirical, method, -condition, -partnersExposure, -ldf_num, -appearance) %>%
  group_by(condition, ldf_num,partnersExposure, appearance,isEmpirical, method) %>%
  summarise(n = n()) %>%
  group_by(condition, ldf_num,partnersExposure, isEmpirical, appearance) %>%
  mutate(n = n/sum(n)) %>%
  ungroup() %>%
  mutate(method = as.factor(method)) %>%
  tidyr::complete(nesting(condition, ldf_num, partnersExposure), isEmpirical, appearance, method, fill = list(n = 0)) %>%
  group_by(condition, partnersExposure, method, isEmpirical, appearance) %>%
  summarise(mean = mean(n), se = seprop(n))


###direct plot of model fit
model_fit <- prop_methods_luce4 %>%
  ungroup() %>%
  mutate(isEmpirical= ifelse(isEmpirical=='method', 'empirical', isEmpirical)) %>%
  unite(mean_new, mean, se) %>%
  spread(isEmpirical, mean_new) %>% 
  separate(empirical, c('mean_emp', 'se_emp'), '_') %>%
  separate(predictions, c('mean_pred', 'se_pred'), '_') %>%
  separate(pred_perf_teaching, c('mean_pred_teach', 'se_pred_teach'), '_') %>%
  # separate(pred_perf_partner, c('mean_pred_partner', 'se_pred_partner'), '_') %>%
  mutate_if(grepl('_', names(.)), as.numeric)


#calculate and plot residuals
residuals <- model_fit %>%
  mutate(residual = (mean_emp -mean_pred_partner)^2) %>%
  group_by(partnersExposure, method, condition, appearance) %>%
  summarise(residual = mean(residual)) %>%
  summarise(residual = sqrt(sum(residual)))

ggplot(residuals, aes(x =as.factor(partnersExposure), y = residual, color = method)) + 
  geom_point()


#plot of model predictions vs empirical game data
model_fit %>%
  rename(mean_pred_plot=mean_pred_teach,
         se_pred_plot=se_pred_teach) %>%
  ggplot(aes(x=mean_pred_plot, y=mean_emp, color=method, shape=as.factor(partnersExposure),
             label = appearance)) +
  geom_point(size=2) +
  geom_abline(intercept =0, slope=1) +
  geom_linerange(aes(ymax = mean_emp + se_emp,
                     ymin = mean_emp - se_emp),
                  position=position_dodge(.25),
                 alpha=.4) +
  geom_errorbarh(aes(xmin = mean_pred_plot - se_pred_plot,
                     xmax = mean_pred_plot + se_pred_plot),
                  position=position_dodge(.25),
                 alpha = .4)+
  # geom_label() +
  #geom_point() + 
  coord_cartesian(xlim=c(0,.8), ylim=c(0,.8)) + 
  labs(y="Empirical Proportion of Trials", x="Model Predictions") +
  scale_color_manual(values=modality.colors, name='Modality', labels=c("Gesture", "Speech", "Teach")) +
  scale_shape_manual(values=c(15,17,8),name="Partner's Exposure", labels=c("None", "Half as Much", "Twice as Much")) 




model_fit %>%
  summarise(cor = cor(mean_emp, mean_pred_teach))

```


# Plot of Referential Communication from LDP Corpus Data
```{r}
# tmp <- getURL("https://raw.githubusercontent.com/dyurovsky/gesture/master/feathers/coded_responses.feather")
library(feather)
corpus_data <- read_feather('empirical_data/coded_responses.feather')


# loaded_subjs<-read_feather('feathers/coded_responses.feather')

modality_data <- corpus_data %>%
  group_by(person, age, subj, freq_cut, rank, freq, referent, modality, chat) %>%
  summarise(n = n()) %>%
  spread(modality, n, fill = 0)


plot_data <- modality_data %>%
  group_by(person, age, subj, freq, referent) %>%
  gather(modality, n, both, gesture, speech) %>%
  mutate(prop = n/sum(n)) %>%
  filter(modality == "both")

#get gradient
cc <- scales::seq_gradient_pal("#87d868", "#124200", "Lab")(seq(0,1,length.out=6))
cc[[6]] <- "#000000"
teaching_ribbons <- scales::seq_gradient_pal("#cbffb7", "#124200", "Lab")(seq(0,1,length.out=6))
teaching_ribbons[[6]] <- "#000000"


ggplot(plot_data, aes(x = log(freq), y = prop, 
                      color = as.factor(age),
                      fill = as.factor(age),
                      label = as.factor(age))) + 
  geom_smooth(method = "loess") + 
  # geom_dl(method = list("first.qp", vjust=c(10.75, 10.5, 9.75, 7.75),hjust=1.05, cex = .8)) +
  # geom_dl(method = list("first.qp", vjust=c(17.5, 17, 14.75, 8),hjust=1.15, cex = .8)) +
  labs(y="Proportion of References\nusing Gesture + Speech", x='Log Referent Frequency') +
  coord_cartesian(ylim=c(0,.4)) +
  # facet_grid(.~age) +
  theme_bw() +
  scale_fill_manual(values=teaching_ribbons, name='Age of Child', labels=c('14 months', '18 months','22 months','26 months', "30 months", "34 months")) +
  scale_color_manual(values=cc, name='Age of Child', labels=c('14 months', '18 months','22 months','26 months', "30 months", "34 months")) 



```




